Investigating Fault-Proneness Through Hotfix Commits:\\A Research in the Light of Different Programming Languages

### 2016-02-17 main.tex
Changed the style of the paper to sig-alternate (ICSE's format) and fixed all the references

\documentclass[conference]{IEEEtran}
\usepackage{blindtext, graphicx}
\usepackage{fancyhdr} 
\graphicspath{ {images/} }
\fancyhf{}
\cfoot{\thepage}
\pagestyle{fancy}

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\pagenumbering{arabic}
\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Investigating Fault-Proneness Through Hotfix Commits: A Research in the Light of Different Programming Languages}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Martin Lazarov\IEEEauthorrefmark{1},
Ismaa'eel Ataullah\IEEEauthorrefmark{2}, Vladyslav Kolesnyk\IEEEauthorrefmark{3},
Veselin Pavlov\IEEEauthorrefmark{4}}
\IEEEauthorblockA{Department of Computer Science,
UCL\\
Email: \IEEEauthorrefmark{1}martin.lazarov.12@ucl.ac.uk,
\IEEEauthorrefmark{2}ismaaeel.ataullah.12@ucl.ac.uk,
\IEEEauthorrefmark{3}vladyslav.kolesnyk.12@ucl.ac.uk,
\IEEEauthorrefmark{4}veselin.pavlov.15@ucl.ac.uk}}

@@

\begin{abstract}
%\boldmath
Choosing the most suitable tools and programming languages is a very challenging task for every project manager and developer. Often the success of a software project is determined by the stability of the system after a release. In this paper, we propose analysing hotfix commits in various development areas as a means of indicating what languages tend to produce the least amount of problems after software has been published. We have focused on parsing metadata from projects hosted on SourceForge related to five specific topics (“software development”, “databases”, “security”, “games”, “front-ends”) and have performed an estimation of the amount of deployed hotfixes to compare the fault-proneness of various languages. 
\end{abstract}

\begin{IEEEkeywords}
Version Control, hotfixes, SourceForge, commits metadata.
\end{IEEEkeywords}

\section{Introduction}
Planning and resource allocation are arguably the most crucial components of any software project. Our research was motivated by the fact that very often, in the planning phase project managers neglect the importance of hotfixes – which becomes essential once a project is deployed. The most common reason for that practice is the fact that hotfixes are not constant and their presence is difficult for prediction. As a result, it is quite challenging to estimate their quantity and incorporate them into the cost of the project.  Our research attempts to provide this crucial information by correlating hotfixes with the choice of language to develop software in a specific area. As there are no clear definitions that specify what a hotfix is, according to our research and observations from the open-source projects, a hotfix can be defined as: “a corrective commit deployed within 1 day, following a project’s version update to fix critical issues in the release”.\par
We believe that the results from our research can provide insights into the average number of issues per project with a focus on different programming topics and the relevant languages used. Consequently, project managers will be able to make a more informed decision with regard to the most appropriate programming languages they could use to try to achieve a minimum number of problems in production. This will improve software engineering practices and the overall project planning activity. Furthermore, it will help optimise the allocation of resources.
The contributions of this paper are:

@@

\section{Related Work}
Many researchers have explored various aspects of defects in software through the use of versioning system history logs. We have also decided to follow this approach due to its openness and accessibility. This method is also very widely adopted amongst the MSR community [1]. Our investigation focused primarily on two research aspects and involved exploring existing work related to: 1) identifying hotfix commits and 2) comparison of programming languages’ stability.

@@

A large number of the reports that we studied implemented a very simple approach. They focused on the commit message as the main mechanism to try to identify whether a commit was of a corrective type [1, 4, 5]. They simply investigated if the string of text contained particular key words, such as: “bug”, “fix”, “issue”, “error”, “broke”, etc.\par

Some authors proposed a solution that involved a slightly more complex procedure including a syntactic analysis of the commit messages. Their algorithm consisted of multiple steps: 1) normalization of the data (removing non-alphanumeric characters), 2) word frequency analysis including a classification of the most common words ( e.g. “for”, “code”, “to”, “the”, etc. ) as being neutral, 3) implementing a keyword clustering algorithm to classify the commits [3].\par

Another approach to identify bug fix commits involved performing a classification based on the size of the commits [1, 3, 5]. The size of a commit can be measured based on the number of lines or the amount of files that were changed. According to these authors, large commits tend to improve code quality or add new features, while small ones are considered to be corrective - targeted at fixing bugs in the project. Some researchers proposed a further fine-grained size classification depending on the number of files changed per commit [1]. They defined the following categories: 1) tiny (1 to 5 changed files), 2) small (6 to 25 files), 3) medium (26 to 125 files), large (126 up ). The scientists further argued that most of the commits tend to be tiny (in 80\% of the cases). This implies that bug fix commits occur very often as corrective commits tend to be small ones.\par

We also looked at the time factor with regard to commits. According to some authors, issues in critical components (e.g. errors in operating system drivers, security errors) tend to be fixed relatively quickly when discovered [2]. In [3], the researchers argue that the time between changes will depend on the type of alteration. Furthermore, they hold the view that corrective changes require the least amount of time to be completed as often they have to be performed promptly regardless of the complexity of the task.\par

In terms of the time aspect, there was another paper which investigated the duration between the moment when a change is applied to a specific branch and the moment the new code was merged into the trunk [10]. However, they explored branches in general without focusing specifically on bugfix branches. Some other scientists were trying to explain people’s behaviour during commits by investigating their commit frequency distribution. As a result, they discovered that on average the time between commits was 3.206 days and half of all the contributors spent less than 13.78 hours between 50\% of their commits [11].\par

Another solution related to identifying defects in software focused on analysing commit metadata to discover design smells and predict where a break in different areas of the code may appear every time a change is performed [6]. However, they did not make their investigation language-specific as opposed to what we propose in our paper.\par

Other scientists tried to explore the logical coupling of commits as a way of identifying potential software defects. They aimed at developing a method to reduce the number of bugs in the code during the development stage so that less fix commits will be needed in the future [9].\par

A further interesting research that we found was related to the use of separate bug fix branches. According to the study performed by Phillips et al. [7], 41\% of the tested projects had incorporated an isolated bug fix branch.

\subsection{Comparison of programming languages' stability}
Throughout our investigation we discovered that there were a lot of authors that tried to utilize commit data to try to propose good engineering practices and find a way to determine code quality [8]. Furthermore, there were several papers that compared and contrasted the performance of different languages, which led to our second research topic. We explored these existing scientific studies to see what techniques authors had implemented. We also analysed the hypotheses these papers proposed to increase the scope of our project.\par

@@

There were researchers that focused on the effect of programming languages on software quality [12]. The authors based performance on the rate of defect occurrences in various languages. The study was carried out on a huge dataset from GitHub (1.5 million commits). To evaluate the effect that programming languages have on software quality, the authors implemented the following strategy. First, they grouped languages according to their type, i.e. strongly vs weakly typed. They also used mathematical methods to triangulate and produce statistics for how group size, project history and number of contributors affect the number of defects. In order to achieve more representative results, they took a sample of 17 languages and extracted the top 50 projects per language. The results showed how defect-prone languages are compared to others. This was done by taking a grand mean and then comparing each language to the grand mean. Languages were then grouped, a mean was taken and then types of languages were once again compared to a grand mean - e.g. functional-strong-static languages compared to functional-dynamic-strong languages. This research focused on classifying bugs based on their type as opposed to our investigation, where we focused specifically on the distribution of hotfixes across programming languages.\par

In [13] Java and C++ were evaluated by comparing bug rates. In this case, there were two definitions for bugs - a defect and a bug itself. A defect included issues such as syntax errors, logical errors and changes in variable name. A bug itself was only defined to be a problem detected during testing or deployment. The mean number of defects and bugs, averaged over a number of projects, was compared between both languages per 1000 lines of code. The scientists also compared observed defects and bugs per hour as well as the time taken to fix defects in both languages. Productivity was also measured by analysing number of lines written per second.\par

There were researchers that used slightly different parameters to compare programming languages, these included program length, programming effort, run-time efficiency, memory consumption, and reliability [14].\par

An interesting paper that we reviewed analysed memory usage and speed of execution of three different bioinformatics methods for six distinct programming languages [15]. This was especially pertinent as bioinformatics applications have huge datasets and computational time taken is not trivial, hence optimising speed is desirable.\par

Another group of authors compared 8 different languages by conciseness, size of executables, running time, memory usage and failure proneness [16]. They argued that conciseness is an important factor as a more compact language can help write a program with fewer bugs.

\section{Solution}
Our hypothesis was motivated by research paper [12]. In it the authors compared the number of bug commits to the total number of commits across 17 languages and 850 different projects.\par

@@

More specifically we wanted to see whether:
\begin{enumerate}
\item \textbf{Strongly-typed languages required fewer hotfixes than weakly-typed languages.}
\item \textbf{Statically-typed languages required fewer hotfixes than dynamically-typed languages.}
\item \textbf{Functional languages required fewer hotfixes than procedural ones.}
\end{enumerate}
\par
In order to evaluate our hypothesis, we developed a tool in Boa [17] that would allow us to incorporate our ideas and perform an analysis to verify them. Next, we present our reasoning and what factors we considered to accomplish our research.\par

@@

Unfortunately, no prior projects implemented a technique to detect version bump commits. However, from the observed data it turned out to be quite straightforward to extract this information by generating a regular expression to match strings like: “bump”, “version”, “update” in an appropriate order.\par

From our Related Work, we found projects [1, 4, 5], which distinguished bug fix commits only based on the commit messages. However, due to the similarity between commit messages for normal bug fixes and hotfix commits, and occasionally the large sequences of text, it was challenging to use the same approach. Therefore, in addition to bug fix commit detection, we applied some other parameters to ensure that it is not just a regular fix, but a hotfix. As previously discussed, we used a regular expression which examined commit messages for words associated with a corrective commit (e.g. “bug”, “fix”, “issue”, “error”, “broke”, etc.) to detect all bug fix commits. Next, we considered other possibilities that would let us detect hotfixes in a more reliable way. Some of the papers we discussed in Related Work [1, 3, 5, 11] incorporated metadata, such as: the size of changes (the number of files that were altered) and the time of the commit (how long it took for a patch to be submitted) to identify bug fixes. We considered these options for our purposes as well. Consequently, in order to increase our confidence level that the current fix commit is an actual hotfix, we introduced the following parameters to verify the emergency of the fix:
\begin{itemize}
  \item \textbf{Time} - commits have to be submitted within 1 day from the version bump commit
  \item \textbf{Look-ahead} range for commits - only n numbers of commits should be considered after version bump
\end{itemize}

@@

Based on our research and the experiments we ran, we managed to identify the most optimal parameters (i.e. \textbf{time}: within 1 day after a release and \textbf{look-ahead}: the first 10 commits - \textit{see Table 1 in Appendix}). Having accomplished that, we focused on selecting a dataset. We decided to analyse SourceForge 2013 because it provided categorisation of projects into programming topics (e.g. code related to: “databases”, “software development”, “security”, etc.). Unfortunately, Github did not possess such classification. We believed that being able to sort projects based on topics would produce greater insights and more useful information for developers, as they would be more interested in language efficiency for particular software development area.\par

@@

\subsection{Limitations.}
There were several limitations that we encountered throughout our research:
\begin{itemize}
  \item we were able to analyse only a small part of all available projects because most of the developers did not support a versioning system, i.e. no bump commits could be identified.
  \item we could not extract hotfix branches, only the overall commits on the trunk were available and thus we used our own parameters to try to distinguish hotfix commits. Therefore, we may have introduced some false positives or failed to account for actual hotfixes.
  \item the way we detected bump commits was by writing a regular expression to match words like:  “bump”, “version”, “update” in an appropriate order. Thus, we may have excluded version updates had they been indicated differently. 
  \item most open source projects are not under the stress of releasing as opposed to commercial projects. Consequently, they do not release very often and/or do not support versioning.
  \item only one dataset was used to evaluate our hypothesis - SourceForge 2013.
\end{itemize}

@@

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

>>>>

\documentclass{sig-alternate}
\usepackage{blindtext, graphicx}
\usepackage{fancyhdr} 
\usepackage{txfonts}
\graphicspath{ {images/} }

\newcommand\FIXME[1]{\textbf{FIXME: #1}}
%\fancyhf{}
%\cfoot{\thepage}
%\pagestyle{fancy}

% *** GRAPHICS RELATED PACKAGES ***
%
%\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
%\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
%\fi

% correct bad hyphenation here
%\hyphenation{op-tical net-works semi-conduc-tor}

%\pagenumbering{arabic}
\begin{document}
\conferenceinfo{MSR}{'16 Austin, Texas USA}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired

\title{Investigating Fault-Proneness Through Hotfix Commits:\\A Research in the Light of Different Programming Languages}
\numberofauthors{4}
\author{
	Martin Lazarov, Ismaa'eel Ataullah, Vladyslav Kolesnyk, Veselin Pavlov\\
	\affaddr{University College London}\\
	\affaddr{London, UK}
}

%\title{Investigating Fault-Proneness Through Hotfix Commits: A Research in the Light of Different Programming Languages}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
%\author{\IEEEauthorblockN{Martin Lazarov\IEEEauthorrefmark{1},
%Ismaa'eel Ataullah\IEEEauthorrefmark{2}, Vladyslav Kolesnyk\IEEEauthorrefmark{3},
%Veselin Pavlov\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{Department of Computer Science,
%UCL\\
%Email: \IEEEauthorrefmark{1}martin.lazarov.12@ucl.ac.uk,
%\IEEEauthorrefmark{2}ismaaeel.ataullah.12@ucl.ac.uk,
%\IEEEauthorrefmark{3}vladyslav.kolesnyk.12@ucl.ac.uk,
%\IEEEauthorrefmark{4}veselin.pavlov.15@ucl.ac.uk}}

@@

\begin{abstract}
%\boldmath
Choosing the most suitable tools and programming languages is a very challenging task for every project manager and developer. Often the success of a software project is determined by the stability of the system after a release. In this paper, we propose analysing hotfix commits in various development areas as a means of indicating what languages tend to produce the least amount of problems after software has been published. We have focused on parsing metadata from projects hosted on SourceForge related to five specific topics (``software development'', ``databases'', ``security'', ``games'', ``front-ends'') and have performed an estimation of the amount of deployed hotfixes to compare the fault-proneness of various languages. 
\end{abstract}

% A category with the (minimum) three required fields
\category{D.2.7}{Software Engineering}{\FIXME{}}
%A category including the fourth, optional field follows...

%\terms{Experimentation, Measurement, Languages, Performance}

\keywords{Version Control, hotfixes, SourceForge, commits metadata}

%\begin{IEEEkeywords}
%Version Control, hotfixes, SourceForge, commits metadata.
%\end{IEEEkeywords}

\section{Introduction}
Planning and resource allocation are arguably the most crucial components of any software project. Our research was motivated by the fact that very often, in the planning phase project managers neglect the importance of hotfixes – which becomes essential once a project is deployed. The most common reason for that practice is the fact that hotfixes are not constant and their presence is difficult for prediction. As a result, it is quite challenging to estimate their quantity and incorporate them into the cost of the project.  Our research attempts to provide this crucial information by correlating hotfixes with the choice of language to develop software in a specific area. As there are no clear definitions that specify what a hotfix is, according to our research and observations from the open-source projects, a hotfix can be defined as: ``a corrective commit deployed within 1 day, following a project's version update to fix critical issues in the release''.\par
We believe that the results from our research can provide insights into the average number of issues per project with a focus on different programming topics and the relevant languages used. Consequently, project managers will be able to make a more informed decision with regard to the most appropriate programming languages they could use to try to achieve a minimum number of problems in production. This will improve software engineering practices and the overall project planning activity. Furthermore, it will help optimise the allocation of resources.
The contributions of this paper are:

@@

\section{Related Work}
Many researchers have explored various aspects of defects in software through the use of versioning system history logs. We have also decided to follow this approach due to its openness and accessibility. This method is also very widely adopted amongst the MSR community \cite{Hattori2008}. Our investigation focused primarily on two research aspects and involved exploring existing work related to: 1) identifying hotfix commits and 2) comparison of programming languages’ stability.

@@

A large number of the reports that we studied implemented a very simple approach. They focused on the commit message as the main mechanism to try to identify whether a commit was of a corrective type~\cite{Hattori2008, Ratzinger2008, Hindle2008}. They simply investigated if the string of text contained particular key words, such as: ``bug'', ``fix'', ``issue'', ``error'', ``broke'', etc.\par

Some authors proposed a solution that involved a slightly more complex procedure including a syntactic analysis of the commit messages. Their algorithm consisted of multiple steps: 1) normalization of the data (removing non-alphanumeric characters), 2) word frequency analysis including a classification of the most common words ( e.g. ``for'', ``code'', ``to'', ``the'', etc. ) as being neutral, 3) implementing a keyword clustering algorithm to classify the commits \cite{Mockus2000}.\par

Another approach to identify bug fix commits involved performing a classification based on the size of the commits \cite{Hattori2008,Mockus2000,Hindle2008}. The size of a commit can be measured based on the number of lines or the amount of files that were changed. According to these authors, large commits tend to improve code quality or add new features, while small ones are considered to be corrective - targeted at fixing bugs in the project. Some researchers proposed a further fine-grained size classification depending on the number of files changed per commit \cite{Hattori2008}. They defined the following categories: 1) tiny (1 to 5 changed files), 2) small (6 to 25 files), 3) medium (26 to 125 files), large (126 up ). The scientists further argued that most of the commits tend to be tiny (in 80\% of the cases). This implies that bug fix commits occur very often as corrective commits tend to be small ones.\par

We also looked at the time factor with regard to commits. According to some authors, issues in critical components (e.g. errors in operating system drivers, security errors) tend to be fixed relatively quickly when discovered \cite{Livshits2005}. In \cite{Mockus2000}, the researchers argue that the time between changes will depend on the type of alteration. Furthermore, they hold the view that corrective changes require the least amount of time to be completed as often they have to be performed promptly regardless of the complexity of the task.\par

In terms of the time aspect, there was another paper which investigated the duration between the moment when a change is applied to a specific branch and the moment the new code was merged into the trunk \cite{Williams2008}. However, they explored branches in general without focusing specifically on bugfix branches. Some other scientists were trying to explain people’s behaviour during commits by investigating their commit frequency distribution. As a result, they discovered that on average the time between commits was 3.206 days and half of all the contributors spent less than 13.78 hours between 50\% of their commits \cite{Kolassa2013}.\par

Another solution related to identifying defects in software focused on analysing commit metadata to discover design smells and predict where a break in different areas of the code may appear every time a change is performed \cite{Oliva2013}. However, they did not make their investigation language-specific as opposed to what we propose in our paper.\par

Other scientists tried to explore the logical coupling of commits as a way of identifying potential software defects. They aimed at developing a method to reduce the number of bugs in the code during the development stage so that less fix commits will be needed in the future \cite{Steff2012}.\par

A further interesting research that we found was related to the use of separate bug fix branches. According to the study performed by Phillips et al. \cite{Phillips2011}, 41\% of the tested projects had incorporated an isolated bug fix branch.

\subsection{Comparison of programming languages' stability}
Throughout our investigation we discovered that there were a lot of authors that tried to utilize commit data to try to propose good engineering practices and find a way to determine code quality \cite{Agrawal2015}. Furthermore, there were several papers that compared and contrasted the performance of different languages, which led to our second research topic. We explored these existing scientific studies to see what techniques authors had implemented. We also analysed the hypotheses these papers proposed to increase the scope of our project.\par

@@

There were researchers that focused on the effect of programming languages on software quality \cite{Ray2014}. The authors based performance on the rate of defect occurrences in various languages. The study was carried out on a huge dataset from GitHub (1.5 million commits). To evaluate the effect that programming languages have on software quality, the authors implemented the following strategy. First, they grouped languages according to their type, i.e. strongly vs weakly typed. They also used mathematical methods to triangulate and produce statistics for how group size, project history and number of contributors affect the number of defects. In order to achieve more representative results, they took a sample of 17 languages and extracted the top 50 projects per language. The results showed how defect-prone languages are compared to others. This was done by taking a grand mean and then comparing each language to the grand mean. Languages were then grouped, a mean was taken and then types of languages were once again compared to a grand mean - e.g. functional-strong-static languages compared to functional-dynamic-strong languages. This research focused on classifying bugs based on their type as opposed to our investigation, where we focused specifically on the distribution of hotfixes across programming languages.\par

In \cite{Phipps1999}, Java and C++ were evaluated by comparing bug rates. In this case, there were two definitions for bugs - a defect and a bug itself. A defect included issues such as syntax errors, logical errors and changes in variable name. A bug itself was only defined to be a problem detected during testing or deployment. The mean number of defects and bugs, averaged over a number of projects, was compared between both languages per 1000 lines of code. The scientists also compared observed defects and bugs per hour as well as the time taken to fix defects in both languages. Productivity was also measured by analysing number of lines written per second.\par

There were researchers that used slightly different parameters to compare programming languages, these included program length, programming effort, run-time efficiency, memory consumption, and reliability \cite{Prechelt2000}.\par

An interesting paper that we reviewed analysed memory usage and speed of execution of three different bioinformatics methods for six distinct programming languages \cite{Nanz2015}. This was especially pertinent as bioinformatics applications have huge datasets and computational time taken is not trivial, hence optimising speed is desirable.\par

Another group of authors compared 8 different languages by conciseness, size of executables, running time, memory usage and failure proneness \cite{Dyer-Nguyen-Rajan-Nguyen-13} \FIXME{IS IT BOA's PAPER HERE?}. They argued that conciseness is an important factor as a more compact language can help write a program with fewer bugs.

\section{Solution}
Our hypothesis was motivated by research paper \cite{Ray2014}. In it the authors compared the number of bug commits to the total number of commits across 17 languages and 850 different projects.\par

@@

More specifically we wanted to see whether:
\begin{enumerate}
\item \textbf{Strongly-typed languages required fewer hotfixes than weakly-typed languages.}
\item \textbf{Statically-typed languages required fewer hotfixes than dynamically-typed languages.}
\item \textbf{Functional languages required fewer hotfixes than procedural ones.}
\end{enumerate}
\par
In order to evaluate our hypothesis, we developed a tool in Boa \cite{Dyer-Nguyen-Rajan-Nguyen-13} that would allow us to incorporate our ideas and perform an analysis to verify them. Next, we present our reasoning and what factors we considered to accomplish our research.\par

@@

Unfortunately, no prior projects implemented a technique to detect version bump commits. However, from the observed data it turned out to be quite straightforward to extract this information by generating a regular expression to match strings like: ``bump'', ``version'', ``update'' in an appropriate order.\par

From our Related Work, we found projects \cite{Hattori2008,Ratzinger2008,Hindle2008}, which distinguished bug fix commits only based on the commit messages. However, due to the similarity between commit messages for normal bug fixes and hotfix commits, and occasionally the large sequences of text, it was challenging to use the same approach. Therefore, in addition to bug fix commit detection, we applied some other parameters to ensure that it is not just a regular fix, but a hotfix. As previously discussed, we used a regular expression which examined commit messages for words associated with a corrective commit (e.g. ``bug'', ``fix'', ``issue'', ``error'', ``broke'', etc.) to detect all bug fix commits. Next, we considered other possibilities that would let us detect hotfixes in a more reliable way. Some of the papers we discussed in Related Work \cite{Hattori2008,Mockus2000,Hindle2008,Kolassa2013} incorporated metadata, such as: the size of changes (the number of files that were altered) and the time of the commit (how long it took for a patch to be submitted) to identify bug fixes. We considered these options for our purposes as well. Consequently, in order to increase our confidence level that the current fix commit is an actual hotfix, we introduced the following parameters to verify the emergency of the fix:
\begin{itemize}
  \item \textbf{Time} - commits have to be submitted within 1 day from the version bump commit
  \item \textbf{Look-ahead} range for commits - only n numbers of commits should be considered after version bump
\end{itemize}

@@

Based on our research and the experiments we ran, we managed to identify the most optimal parameters (i.e. \textbf{time}: within 1 day after a release and \textbf{look-ahead}: the first 10 commits - \textit{see Table 1 in Appendix}). Having accomplished that, we focused on selecting a dataset. We decided to analyse SourceForge 2013 because it provided categorisation of projects into programming topics (e.g. code related to: ``databases'', ``software development'', ``security'', etc.). Unfortunately, Github did not possess such classification. We believed that being able to sort projects based on topics would produce greater insights and more useful information for developers, as they would be more interested in language efficiency for particular software development area.\par

@@

\subsection{Limitations.}
There were several limitations that we encountered throughout our research:
\begin{itemize}
  \item we were able to analyse only a small part of all available projects because most of the developers did not support a versioning system, i.e. no bump commits could be identified.
  \item we could not extract hotfix branches, only the overall commits on the trunk were available and thus we used our own parameters to try to distinguish hotfix commits. Therefore, we may have introduced some false positives or failed to account for actual hotfixes.
  \item the way we detected bump commits was by writing a regular expression to match words like:  ``bump'', ``version'', ``update'' in an appropriate order. Thus, we may have excluded version updates had they been indicated differently. 
  \item most open source projects are not under the stress of releasing as opposed to commercial projects. Consequently, they do not release very often and/or do not support versioning.
  \item only one dataset was used to evaluate our hypothesis - SourceForge 2013.
\end{itemize}

@@

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
%\ifCLASSOPTIONcaptionsoff
%  \newpage
%\fi
